Effects of the body force on the evacuation dynamics

### 2020-01-01 paper.tex
algunas correcciones de gramatica

In this section, we present the results corresponding to the bottleneck geometry. We show how the overall dynamic of the system changes depending on the $k$ value associated to the body force in the original SFM.\\

Fig. \ref{vd_vs_te} shows the evacuation time as a function of the pedestrian's desired velocity. The evacuation time was defined as the time until 70\% of the pedestrians left the room. The different markers correspond to different $k$ values (\textit{i.e.} different body ``stiffness''). The crosses correspond to the Helbing original SFM parameter, the up-triangles correspond to the value measured in Ref ..., squares correspond to an absent body force and circles correspond to an extreme value of stiffness (one order of magnitude higher than the original SFM). The down triangles correspond to an intermediate value between the empirical value presented in Ref... and the chosen by Helbing in Ref...\\

There is a noticeable difference in the evacuation time depending on the $k$ value of the body force. The higher the $k$ the smaller the evacuation time, meaning that if pedestrians are stiffer, the overall speed increases (hence the evacuation time decreases).\\

The Faster-is-Faster (FIF) is the phenomenon in which under certain conditions, increasing the desired velocity yields to a reduction of the evacuation time. The Faster-is-Faster is the opposite effect (the higher $v_d$ the lower the evacuation time).  Notice that the presence of the Faster-is-Faster (FIF) effect and the Faster-is-Slower are dependent on the value of $k$. In Fig.~\ref{vd_vs_te} we can see that $k=1.2$~E5 only exhibits FIS effect for $v_d>$2~m/s. When $k=1.2$~E6, the evacuation time reduces as $v_d$ increases, meaning that only FIF is reported. The explored values of $k$ that satisfy $k \leq 6$~E4 exhibit both the FIS and the FIF effect depending on the range of desired velocities. Notice that the behavior of the case in which $k = 0$ is very similar to the case in which $k = 1.2$~E4.\\
% se puede discutir que este resultado es importante porque k=1.2E4 es un valor empirico y k=1.2E5 tambien lo es, osea que el FIF podria estar o no estar dependiendo de cual es el verdadero valor de k. Habria que chequear en unidades reducidas como queda.

@@

Network analysis is a set of techniques that characterize the structures in a graph. We defined the graph of contacts among individuals. In this graph, each node represents a pedestrian and the link between a pair of nodes is settled if the pedestrians are in physical contact (\textit{i.e.} $r_{ij}<R{ij}$).\\

Fig.~\ref{degree_vd} shows the mean degree of the contact graph as a function of the desired velocity. The degree of a node is defined as the amount of edges assigned to the node. This means, the number of pedestrians that are in physical contact with a given pedestrian. The mean degree is the average of the degree over all the nodes (pedestrians) and over the whole sampled time (from the time the system reaches the stationary state to the end of the simulation).\\

Increasing $v_d$ increases the mean degree in all the cases. This happens because increasing $v_d$ produces higher densities that force individuals to touch each other. For a given $v_d$ the mean degree reduces as the $k$ value increases. In other words, the stiffer the pedestrians the less they touch each other. We will later discuss that this phenomenon is restricted to the bottleneck geometry that cannot be extrapolated to the corridor geometry.\\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./degree_vs_vd_multi_kn.eps}\label{degree_vd}}\ 
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./overlap_vs_vd_multi_kn.eps}\label{overlap_vd}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the pedestrian’s desired velocity (m/s). (b) Mean overlap as a function of the pedestrian’s desired velocity (m/s). Each symbol indicates the $k$ value corresponding to the body force (see the labels). The data corresponds to a bottleneck with periodic boundary conditions (re-injecting pedestrians). The average was taken every two seconds once the crowd reaches the stationary state ($t=$20~s) until the end of the simulation ($t=$110~s).}
\label{degree_overlap_vd}
\end{figure}

@@

Comparing Fig. \ref{network_vd2_kn0} against Fig. \ref{network_vd2_kn1200000} and Fig \ref{network_vd10_kn0} against Fig. \ref{network_vd10_kn1200000} it is clear that increasing the stiffness reduces the degree. The most notorious case is the pictures corresponding to $v_d=$10 in which, despite having the same $v_d$, the size of the bulk increases when the pedestrians are stiffer. Moreover, when pedestrians are stiffer they have fewer contacts in general (less degree). Notice that there are no pedestrians with degree 6 when $k=1.2$~E6 and there are a lot of pedestrians with degree 6 when $k$=0.

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./network_vd2_kn0.eps}\label{network_vd2_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./network_vd10_kn0.eps}\label{network_vd10_kn0}}\\
        \subfloat[]{\includegraphics[width=0.45\columnwidth]{./network_vd2_kn1200000.eps}\label{network_vd2_kn1200000}}\ 
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./network_vd10_kn1200000.eps}\label{network_vd10_kn1200000}}\\
\caption[width=0.47\columnwidth]{Contact networks of a 225 pedestrian evacuation through a bottleneck. The door is placed at $(x,y)=(20,10)$~m, the width of the door is 0.92~m (equivalent to 2 pedestrian's diameter). The lines that connect the nodes (pedestrians) represent the contact between them. The color represents the degree (the number of pedestrians with which it is connected). \ref{network_vd2_kn0} and \ref{network_vd10_kn0} correspond to a simulation without body force with $v_d=$2 and $v_d=$10 respectively. \ref{network_vd2_kn1200000} and \ref{network_vd10_kn1200000} correspond to simulations with $k$=1.2~E6 with $v_d=$2 and $v_d=$10 respectively.}
\label{network_bottleneck}
\end{figure}

@@

In this section, we present the results corresponding to the corridor geometry. We show how the overall dynamic of the system changes depending on the $k$ value associated to the body force in the original SFM.\\

In Fig. \ref{flow_density} we show the flow as a function of the global density (fundamental diagram). \ref{flow_density_no_body_force} corresponds to non body force (\textit{i.e.} $k_n=0$), \label{flow_density_body_force} corresponds to a body force with $k_n$=1.2~E5. Each curve correspond to a different friction value. The circles correspond to the value of the original SFM ($\kappa = 2.4$~E5), the "+" symbol correspond to the original value increased by a factor of five and triangles correspond to the original value increased by a factor of ten. \\

According to most of the empirical measurements, the flow decreases for high enough values of density because the motion becomes jammed. Notice that the original friction does not produce a reduction of the flow when there is no body force but when the body force is included (with the original value), the flow reduces from  5 p m$^{-2}$ to 7 p m$^{-2}$ and increases for higher values of density. When the density is five times the original value, the flow reduces for densities higher than 5 p m$^{-2}$ in both cases (with body force and without body force). The main qualitative difference is that including the body force produces a subtle increment in the flow for densities higher than 7 p m$^{-2}$. When the friction is increased by a factor of ten, the flow reduces or remains constant when the density is higher than 5 p m$^{-2}$ for both cases (with body force and without body force). This result suggests that even including the body force, the original SFM requires an increment of the friction in order to reproduce the flow reduction reported in empirical measurements.\\

It is worth mentioning that increasing the body force reduces the flow, for a given density and a given friction value (compare \ref{flow_density_no_body_force} with \ref{flow_density_body_force} ). In other words, pedestrians walk slower along the corridor if the $k$ value increases (\textit{i.e.} if they are stiffer). This result is the opposite of what we get in bottleneck geometry since the stiffer the pedestrians the faster they walk.\\

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./flow-density_multifric_nobodyforce.eps}\label{flow_density_no_body_force}}\ 
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./flow-density_multifric_bodyforce.eps}\label{flow_density_body_force}}\\
\caption[width=0.47\columnwidth]{Flow vs. global density. The flow is calculated in a circular area of R=1~m at the center of the corridor. The circle corresponds to the original friction of the SFM, the "+" symbol corresponds to the friction increased by a factor of five and the triangles correspond to the friction increased by a factor of ten. The desired velocity was $v_d=1$ in all the cases. Fig. \ref{flow_density_no_body_force} corresponds to simulations without body force and Fig. \ref{flow_density_body_force} corresponds to a body force with $k=$1.2~E5.}
\label{flow_density}
\end{figure}

To understand why in corridor geometry pedestrians walk slower as they become stiffer it is important to take into account that pedestrians walk slower if the friction increases. In corridors, the friction increases when pedestrians become stiffer because the crowd tends to cluster more. The clusterization means that more pedestrians are touching each other. \\

One way to analyze the clusterization is by creating a network out of the (x,y) positions of the pedestrians. In this network, each node represents a pedestrian and there is a link between a pair of nodes $i,j$ if the pair of pedestrians represented by the nodes are touching each other (\textit{i.e.} $r_{ij} \leq R_{ij}$). Once the network is created, it is possible to use the analytical tools that allow describing the configuration such as the mean degree, the distribution of clusters, the shortest path length, etc.\\

@@

It is important to notice that in Fig. \ref{degree_dens} for a given value of density, the stiffer the pedestrians, the higher the degree. A higher degree means more friction force which is the cause of the reduction of the flow. In the bottleneck geometry, the opposite thing happens, the stiffer the pedestrians the lower the mean degree. This is the reason why in bottleneck pedestrians evacuate faster as they become stiffer while in corridor pedestrian's flow is lower as they become stiffer.\\

>>>>

In this section, we present the results corresponding to the bottleneck geometry. We show the effects of modifying the body force coefficient $k$ in the evacuation dynamics. This coefficient is associated with the stiffness of the pedestrians. \\

Fig. \ref{vd_vs_te} shows the evacuation time as a function of the pedestrian\textsc{\char13}s desired velocity. The evacuation time was defined as the time until 80\% of the pedestrians left the room. The different markers correspond to different $k$ values (\textit{i.e.} different body ``stiffness''). The crosses correspond to the Helbing original SFM parameter, the up-triangles correspond to the value measured in Ref ..., squares correspond to zero body force and circles correspond to an extreme value of stiffness (one order of magnitude higher than the original SFM). The down triangles correspond to an intermediate value between the empirical value presented in Ref... and the chosen by Helbing in Ref...\\

There is a noticeable difference in the evacuation time depending on the $k$ value of the body force. The higher is $k$ the smaller the evacuation time, meaning that if pedestrians are stiffer, the overall speed increases (hence the evacuation time decreases).\\

The Faster-is-Faster (FIF) is the phenomenon in which under certain conditions, increasing the desired velocity yields to a reduction of the evacuation time. The Faster-is-Slower is the opposite effect (the higher $v_d$, the higher the evacuation time).  Notice that the presence of the Faster-is-Faster (FIF) effect and the Faster-is-Slower are dependent on the value of $k$. In Fig.~\ref{vd_vs_te} we can see that $k=1.2$~E5 only exhibits FIS effect for $v_d>$2~m/s. When $k=1.2$~E6, the evacuation time reduces as $v_d$ increases, meaning that only FIF is reported. The explored values of $k$ that satisfy $k \leq 6$~E4 exhibit both the FIS and the FIF effect depending on the range of desired velocities. Notice that the behavior of the case in which $k = 0$ is very similar to the case in which $k = 1.2$~E4.\\
% se puede discutir que este resultado es importante porque k=1.2E4 es un valor empirico y k=1.2E5 tambien lo es, osea que el FIF podria estar o no estar dependiendo de cual es el verdadero valor de k. Habria que chequear en unidades reducidas como queda.

@@

Network analysis is a set of techniques that characterize the structures in a graph. We defined the contact network where each node represents a pedestrian and the link between a pair of nodes is settled if the pedestrians are in physical contact (\textit{i.e.} $r_{ij} \leq R{ij}$).\\

Fig.~\ref{degree_vd} shows the mean degree of the contact network as a function of the desired velocity. The degree of a node is defined as the amount of edges assigned to the node. This means, the number of pedestrians that are in physical contact with a given pedestrian. The mean degree is the average of the degree over all the nodes (pedestrians) and over the whole sampled time (from the time the system reaches the stationary state to the end of the simulation).\\

Increasing $v_d$ increases the mean degree in all the cases. This happens because increasing $v_d$ produces higher densities that force individuals to touch each other. For a given $v_d$ the mean degree reduces as the $k$ value increases. In other words, the stiffer the pedestrians are, the less they touch each other. We will later discuss that this phenomenon is restricted to the bottleneck geometry that cannot be extrapolated to the corridor geometry.\\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./degree_vs_vd_multi_kn.eps}\label{degree_vd}}\ 
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./overlap_vs_vd_multi_kn.eps}\label{overlap_vd}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the pedestrian’s desired velocity (m/s). (b) Mean overlap as a function of the pedestrians desired velocity. Each symbol indicates the $k$ value corresponding to the body force (see the labels). The data corresponds to a bottleneck with periodic boundary conditions (re-injecting pedestrians). The average was taken every two seconds once the crowd reaches the stationary state ($t=$20~s) until the end of the simulation ($t=$110~s).}
\label{degree_overlap_vd}
\end{figure}

@@

Comparing Fig. \ref{network_vd2_kn0} against Fig. \ref{network_vd2_kn1200000} and Fig \ref{network_vd10_kn0} against Fig. \ref{network_vd10_kn1200000} it is clear that increasing the stiffness reduces the degree. The most notorious case is exhibit in the pictures corresponding to $v_d=$10 in which, despite having the same $v_d$, the size of the bulk increases when the pedestrians are stiffer. Moreover, when pedestrians are stiffer they have fewer contacts in general (less degree). Notice that there are no pedestrians with degree 6 when $k=1.2$~E6 and there are a lot of pedestrians with degree 6 when $k$=0.

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./network_vd2_kn0.eps}\label{network_vd2_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./network_vd10_kn0.eps}\label{network_vd10_kn0}}\\
        \subfloat[]{\includegraphics[width=0.45\columnwidth]{./network_vd2_kn1200000.eps}\label{network_vd2_kn1200000}}\ 
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./network_vd10_kn1200000.eps}\label{network_vd10_kn1200000}}\\
\caption[width=0.47\columnwidth]{Contact networks of a 225 pedestrian evacuation through a bottleneck. The door is placed at $(x,y)=(20,10)$~m, the width of the door is 0.92~m (equivalent to 2 pedestrian\textsc{\char13}s diameter). The lines that connect the nodes (pedestrians) represent the contact between them. The color represents the degree (the number of pedestrians with which it is connected). \ref{network_vd2_kn0} and \ref{network_vd10_kn0} correspond to a simulation without body force with $v_d=$2 and $v_d=$10 respectively. \ref{network_vd2_kn1200000} and \ref{network_vd10_kn1200000} correspond to simulations with $k$=1.2~E6 with $v_d=$2 and $v_d=$10 respectively.}
\label{network_bottleneck}
\end{figure}

@@

In this section, we present the results corresponding to the corridor geometry. We show the effects of modifying the body force coefficient $k$ in the dynamics.\\

In Fig. \ref{flow_density} we show the flow as a function of the global density (fundamental diagram). \ref{flow_density_no_body_force} corresponds to zero body force (\textit{i.e.} $k_n=0$), \ref{flow_density_body_force} corresponds to a body force with $k_n$=1.2~E5. Each curve correspond to a different friction value. The circles correspond to the value of the original SFM ($\kappa = 2.4$~E5), the "+" symbol correspond to the original value increased by a factor of five and triangles correspond to the original value increased by a factor of ten. \\

According to most of the empirical measurements, the flow decreases for high enough values of density because pedestrians become jammed. Notice that the original friction does not produce a reduction of the flow when there is zero body force but when the body force is included (with the original SFM value), the flow reduces from  5 p m$^{-2}$ to 7 p m$^{-2}$ and increases for higher values of density. When the friction is five times the original value, the flow reduces for densities higher than 5 p m$^{-2}$ in both cases (with body force and without body force). The main qualitative difference is that including the body force produces a subtle increment in the flow for densities higher than 7 p m$^{-2}$. When the friction is increased by a factor of ten, the flow reduces or remains constant when the density is higher than 5 p m$^{-2}$ for both cases (with body force and without body force). This result suggests that even including the body force, the original SFM requires an increment of the friction in order to reproduce the flow reduction reported in empirical measurements.\\

It is worth mentioning that increasing the body force reduces the flow, for a given density and a given friction value (compare \ref{flow_density_no_body_force} with \ref{flow_density_body_force} ). In other words, pedestrians walk slower along the corridor if the $k$ value increases (\textit{i.e.} if they are stiffer). This result is the opposite of what we get in bottleneck geometry since in bottleneck, the stiffer the pedestrians are, the faster they walk (and evacuate).\\

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./flow-density_multifric_nobodyforce.eps}\label{flow_density_no_body_force}}\ 
    \subfloat[]{\includegraphics[width=0.45\columnwidth]{./flow-density_multifric_bodyforce.eps}\label{flow_density_body_force}}\\
\caption[width=0.47\columnwidth]{Flow vs. global density. The flow is calculated in a circular area of R=1~m at the center of the corridor. The circular markers correspond to the original friction of the SFM, the "+" symbol corresponds to the friction increased by a factor of five and the triangles correspond to the friction increased by a factor of ten. The desired velocity was $v_d=1$ in all the cases. Fig.~\ref{flow_density_no_body_force} corresponds to simulations without body force and Fig. \ref{flow_density_body_force} corresponds to a body force with $k=$1.2~E5.}
\label{flow_density}
\end{figure}

To understand why in corridor geometry pedestrians walk slower as they become stiffer it is important to take into account that pedestrians walk slower if the friction increases. In corridors, the friction increases when pedestrians become stiffer because the crowd tends to be more connected. To be more connected means that more pedestrians are touching each other. \\

One way to analyze the increment of connectivity is by creating a network out of the (x,y) positions of the pedestrians. In this network, each node represents a pedestrian and there is a link between a pair of nodes $i,j$ if the pair of pedestrians represented by the nodes are touching each other (\textit{i.e.} $r_{ij} \leq R_{ij}$). Once the network is created, it is possible to use the analytical tools that allow describing the network such as the mean degree, the distribution of clusters, the shortest path length, etc.\\

@@

It is important to notice that in Fig. \ref{degree_dens} for a given value of density, the stiffer the pedestrians, the higher the degree. A higher degree means more friction force which is the cause of the reduction of the flow. In the bottleneck geometry, the opposite thing happens, the stiffer the pedestrians the lower the mean degree. This is the reason why in bottleneck pedestrians evacuate faster as they become stiffer while in corridor pedestrian\textsc{\char13}s flow is lower as they become stiffer.\\
